babette: BEAUti 2, BEAST2 and Tracer for R

### 2018-02-15 article.tex
Feedback @Giappo

\verb;babette; is written in the R programming language \cite{R}
and enables the full BEAST2 workflow from an R function call,
in a similar way that BEAUti, DensiTree and Tracer do.

@@

\verb;babette; currently has 61 exported functions to set up  
a BEAST2 configuration file. 
\verb;babette; is an alternative for a majority of BEAUti use cases, 
but does not yet support the full functionality of BEAUti. 
Because of BEAUti's high number of plugins, 
\verb;babette; uses a software architecture that expects te be extended.

\verb;babette; has 7 exported function to run and help run BEAST2.
One function is a wrapper function to run BEAST2, others
allow the user to check if a BEAST2 configuration file is indeed valid.

@@

This code will create a (temporary) BEAST2 configuration file,
using a FASTA file with name \verb;alignment.fas;, 
using the same default settings as BEAUti.
\verb;babette; will then execute a BEAST2 with that file, and
parses the output. The output, named \verb;out;, is a list of
parameter estimates, posterior phylogenies (one per alignment)
and operator acceptances.

@@

\verb;babette; also uses the same default distributions as BEAUti 
for the site models, clock models and tree priors. 
For example, a Yule tree prior assumes the birth rate follows a uniform distribution, 
from minus infinity to plus infinity. 
This assumption implies that negative and positive birth rates are just as likely, 
where a negative birth rate is biologically impossible (note that 
in practice, this usually works out just fine).
One may prefer an exponential distribution instead, 
as this would assume only positive birth rates, 
and makes high birth rates unlikely.

@@

\verb;babette; allows for the same functionality of Tracer. 
Tracer, among others, shows the effective sample
sizes of each posterior's variables. As an MCMC run needs to
converge first, Tracer discards the first 10\% of all parameter
estimates. This is called the burn-in. To calculate the effective
sample sizes in \verb;babette;, with the same burn-in of 10\%:

\begin{lstlisting}[language=R, floatplacement=H]
traces <- remove_burn_ins(traces = out$estimates, burn_in_fraction = 0.1)
esses <- calc_esses(traces, sample_interval = 1000)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, floatplacement=H]
sum_stats <- calc_sum_stats(
  out$estimates$birthRate, 
  sample_interval = 1000, 
  burn_in_fraction = 0.1
)
\end{lstlisting}

>>>>

\verb;babette; is written in the R programming language \cite{R}
and enables the full BEAST2 workflow from an R function call,
in a similar way to what BEAUti, DensiTree and Tracer do.

@@

\verb;babette; currently has 61 exported functions to set up  
a BEAST2 configuration file. 
\verb;babette; is an alternative for a majority of BEAUti use cases, 
but does not yet support the full functionality of BEAUti. 
Because of BEAUti's high number of plugins, 
\verb;babette; uses a software architecture that expects to be extended.

\verb;babette; has 7 exported function to run and help run BEAST2.
One function is used to run BEAST2, others
allow the user to check if a BEAST2 configuration file is indeed valid.

@@

This code will create a (temporary) BEAST2 configuration file,
using a FASTA file with name \verb;alignment.fas;, 
using the same default settings as BEAUti.
\verb;babette; will then execute BEAST2 with that file, and
parses the output. The output, named \verb;out;, is a list of
parameter estimates, posterior phylogenies (one for each alignment)
and operator acceptances.

@@

\verb;babette; also uses the same default prior distributions as BEAUti 
for each of the site models, clock models and tree priors. 
For example, by default, a Yule tree prior assumes the birth rate 
follows a uniform distribution, 
from minus infinity to plus infinity. 
This assumption implies that negative and positive birth rates are just as likely, 
where a negative birth rate is biologically impossible (note that 
in practice, this usually works out just fine).
One may prefer an exponential distribution instead, 
as this would assume only positive birth rates, 
and makes high birth rates unlikely.

@@

\verb;babette; allows for the same functionality of Tracer. 
Tracer, among others, shows the effective sample
sizes of each posterior's variables. As an MCMC run needs to
converge first, Tracer discards the first 10\% of all parameter
estimates. This is called the "burn-in". To remove the burn-in
in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
traces <- remove_burn_ins(out$estimates)
\end{lstlisting}

Then, to calculate the effective sample sizes in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
esses <- calc_esses(
  traces, 
  sample_interval = 1000
)
\end{lstlisting}

@@

\begin{lstlisting}[language=R, floatplacement=H]
sum_stats <- calc_summary_stats(
  traces$birthRate, 
  sample_interval = 1000
)
\end{lstlisting}
